Supplement for: Carbon cycling in mature and regrowth forests globally

### 2021-01-25 manuscript/ERL_global_C_review.Rmd
produce doc with highlighted changes, fix typos in submission version and .Rmd
#67

*Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and boreal. We calculaate statistics for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.

@@

We grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
Climate zones (Fig. `r figure_map`) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
We defined the tropical biome as including all equatorial (A) zones, temperate biomes as including all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and the boreal biome as including the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
We defined leaf type (broadleaf / needleleaf) based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics.
We also classified forests as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
Assigning stands to these groupings required the exclusion of records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database) or records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database). 
We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
In total, our analysis retained `n_records` records. 
Numbers of records by biome and age class are given in Table S1.

Data were summarized to produce schematics of C cycling for mature forets of each biome.
To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
Values were then averaged across geographically distinct areas, defined as plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
This step was taken to avoid pseudo-replication.

@@

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood. In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction. The boxplot illustrates distribution across mature forests, with different letters indicating signifant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each stock with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs. S20-S30).", out.width='100%', fig.pos='H'}

>>>>

*Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we use `r n_records` *ForC* records for `r nrow(T1)` C cycle variables from `r n_areas_tot` geographic locations to characterize ensemble C budgets for four broad forest types -- tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and boreal. We calculate statistics for both mature and regrowth (age <100 years) forests, and quantify trends with stand age in regrowth forests for all variables with sufficient data.

@@

We grouped forests into four broad biome types based on climate zones and dominant vegetation type (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature).
Climate zones (Fig. `r figure_map`) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
We defined the tropical biome as including all equatorial (A) zones, temperate biomes as including all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and the boreal biome as including the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd).
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis.
We defined leaf type (broadleaf / needleleaf) based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006].
For young tropical forests imported from *GROA* but not yet classified by leaf type, we assumed that all were broadleaf, consistent with the rarity of naturally regenerating needleleaf forests in the tropics.
We also classified forests as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication).
Assigning stands to these groupings required the exclusion of records for which *ForC* lacked geographic coordinates (`r round(n_sites_coord_unk/nrow(ForC_sites)*100, digits=1)`% of sites in full database) or records of stand age (`r round(n_records_age_unk/nrow(ForC_measurements)*100, digits=1)`% of records in full database). 
We also excluded records of stand age = 0 year (`r round(n_records_age_0/nrow(ForC_measurements)*100, digits=1)`% of records in full database).
In total, our analysis retained `r n_records` records. 
Numbers of records by biome and age class are given in Table S1.

Data were summarized to produce schematics of C cycling for mature forests of each biome.
To obtain the values reported in the C cycle schematics, we first averaged any repeated measurements within a plot.
Values were then averaged across geographically distinct areas, defined as plots clustered within 25 km of one another [*sensu* @anderson-teixeira_forc_2018], weighting by area sampled if available for all records.
This step was taken to avoid pseudo-replication.

@@

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap = "Figure 8 | Age trends and biome differences in some of the major forest C stocks: (a) aboveground biomass, (b) foliage, (c) fine roots, (d) dead wood. In each panel, the left scatterplot shows age trends in forests up to 100 years old, as characterized by a linear mixed effects model with fixed effects of log10(age) and biome. The fitted line indicates the effect of age (solid lines: significant at p<0.05, dashed lines: non-significant), and non-parallel lines indicate a significant log10(age) x biome interaction. The boxplot illustrates distribution across mature forests, with different letters indicating significant differences between biomes. Data from biomes that did not meet the sample size criteria (see Methods) are plotted, but lack regression lines (young forests) or test of differences across biomes (mature forests). Individual figures for each stock with sufficient data, along with maps showing geographic distribution of the data, are given in the Supplement (Figs. S20-S30).", out.width='100%', fig.pos='H'}
